---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(ggplot2)
library(dplyr)
unzip("activity.zip")
x=read.csv("activity.csv")
x$date=(as.Date(as.character(x$date),"%Y-%m-%d"))
```

## What is mean total number of steps taken per day?
Sum of steps per day:
```{r}
stepdate=aggregate(steps~date,x,sum,na.rm=T)
stepdate
```

Histogram of the above data:
```{r}
ggplot(stepdate,aes(steps))+geom_histogram(binwidth=700)+ggtitle("Total Number of Steps Taken Each Day")
```

The mean of total number of steps taken per day is:
```{r}
meanstep=mean(stepdate$steps)
meanstep
```

The median of total number of steps taken per day is:
```{r}
medstep=median(stepdate$steps)
medstep
```

## What is the average daily activity pattern?
5-minute interval by the average number of steps taken, averaged across all days 
```{r}
stepmeanint=aggregate(steps~interval,x,mean,na.rm=T)
ggplot(stepmeanint,aes(interval,steps))+geom_line()+ggtitle("Average Number of Steps VS 5-minute intervals")

```

Interval containing maximum average number of steps is 
```{r}
maxintrow=which(stepmeanint$steps==max(stepmeanint$steps))
stepmeanint[maxintrow,1]
```

this is the whole row
```{r}
stepmeanint[maxintrow,]
```
## Imputing missing values
This is the number of NAs:
```{r}
seena=table(is.na(x))
seena[2]
```

to get rid of the NAs, we will find and replace missing values in steps, we'll use the total mean steps from above for filler, what could possibly go wrong?
```{r}
#subset the NA so we can pinpoint the missing data
nadata=subset(x,is.na(x))
#isolate the ones that aren't missing too
notnadata=subset(x,!is.na(x))
nadateslist=as.list(unique(nadata$date))
#fill in the missing data with previous mean
nadata$steps=meanstep
#create new dataset with no more missing data
full=rbind(nadata,notnadata)
fullstepdate=aggregate(steps~date,full,sum,na.rm=T)
#create histogram
ggplot(full,aes(steps))+geom_histogram(binwidth=2000)+ggtitle("Total Number of Steps Taken Each Day")
```

The mean of total number of steps taken per day after adding the filler becomes:
```{r}
fullmeanstep=mean(fullstepdate$steps)
fullmeanstep
```

The mean of total number of steps taken per day after adding the filler becomes:
```{r}
fullmedstep=median(fullstepdate$steps)
fullmedstep
```

As we can see above, imputing the NA -- especially in this case where there were 2304 NAs of the total of 17568 observations -- could cause major change in the analysis.


## Are there differences in activity patterns between weekdays and weekends?
add to the dataset actual days 
```{r}
xdays=mutate(x,days=weekdays(date))
```

categorize weekdays and weekends
```{r}
xdays=xdays %>% mutate(week = ifelse(xdays$days == "Monday","weekday",ifelse(xdays$days == "Tuesday","weekday",ifelse(xdays$days =="Wednesday","weekday",ifelse(xdays$days =="Thursday","weekday",ifelse(xdays$days =="Friday","weekday",ifelse(xdays$days == "Saturday","weekend",ifelse(xdays$days =="Sunday","weekend", 0))))))))
```

plot average number of steps taken VS weekday days or weekend days:
```{r}
stepmeanintweek=aggregate(steps~interval+week,xdays,mean,na.rm=T)
ggplot(stepmeanintweek,aes(interval,steps))+geom_line()+ggtitle("Average Number of Steps VS 5-minute intervals")+facet_grid(.~week)
```

people seem to be more active during weekends.